---
title: "Data Import"
output: github_document
---

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(haven)
```


I'm an R Markdown document! 

## Data Import: CSVs

Let's import data using the 'readr' package

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
`janitor::clean_names` is used to clean the special characters in the column names in order to prevent bugs in further data loading


Look at the data

```{r}
litters_df
head(litters_df)
tail(litters_df)
```


```{r, eval = FALSE}
View(litters_df)
```

```{r}
skimr::skim(litters_df)
```

`read..csv` options ..

```{r}
read_csv("data/FAS_litters.csv", na = c("","NA", 999, 88), skip = 2)
```
That means you could manually set the NA-defined value.
`skip` allows you to skip the first several rows when the author added useless info into data.


## Other file formats

We need to read in an excel spreadsheet
```{r}
mlb_df = read_excel("data/mlb11.xlsx")
```

```{r}
lotr_words_df = 
  read_excel(
    "data/LotR_Words.xlsx",
    range = "B3:D6"
  )
```

## Still more formats

Read in a SAS dataset.

Remember not to commit or push a temp data file appeared on the rightside. Close it and then commit or push the files.
```{r}
pulse_df =read_sas("data/public_pulse_data.sas7bdat")
```


## Data Export
```{r}
write_csv(lotr_words_df, "data/lotr_words_df.csv")
```


## Why not base R??

```{r}
dont_do_this_df = read.csv("data/FAS_litters.csv")
```

if do not want to update the data becauses its privacy or size, type in `.gitignore` and add the data source address.